A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-09 Stickleback_Paper.tex
.

An alternative explanation is that at lower rates of gene flow,
freshwater alleles present as standing variation in the marine habitat
are more tightly linked to marine alleles,
and so adaptation in the new lakes must wait for recombination.
This might be expected since freshwater alleles present in the ocean,
to remain present at migration-selection balance at low migration rates,
must be ``masked'' by nearby compensatory alleles.
Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
Said another way,
higher gene flow from freshwater to the ocean
maintains relatively intact freshwater haplotypes
that can be more easily rebuilt.
To confirm this, we calculated for each individual in the initial generation of the new lakes
the total contribution to trait values of each of the ten genomic regions 
that containing trait-affecting mutations.
The recombination rate within each of these ten regions is $10^{-2}$,
so these regions act initially more-or-less like single loci.
If a region moves the trait an average of $x$ units towards the freshwater optimum
(averaging over dominance relationships),
Given the probability that an ``effect region'' fixes at a given trait value, 
$2 \times x / 45$ where $x$ is the effect size of a haplotype,
we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
Below, we see the mean of these expected total effect sizes across all lakes.

>>>>

An alternative explanation is that at lower rates of gene flow,
freshwater alleles present as standing variation in the marine habitat
are more tightly linked to marine alleles,
and so adaptation in the new lakes must wait for recombination.
This might be expected since freshwater alleles present in the ocean,
to remain present at migration-selection balance at low migration rates,
must be ``masked'' by nearby compensatory alleles.
Reversing this ``masking'' then slows the process of adaptation that uses this genetic variation.
Said another way,
higher gene flow from freshwater to the ocean
maintains relatively intact freshwater haplotypes
that can be more easily rebuilt.
To confirm this, we calculated for each individual in the initial generation of the new lakes
the total contribution to trait values of each of the ten genomic regions 
that containing trait-affecting mutations.
The recombination rate within each of these ten regions is $10^{-2}$,
so these regions act initially more-or-less like single loci.
Even though the trait is additive across loci,
the presence of sufficiently many distinct alleles to attain, in principle,
the optimal freshwater trait value
does not ensure adaptation will actually occur with these alleles.
On the contrary, many of these will likely be lost to drift.
To get a rough idea of whether the genetic variation present in the initial population
is sufficient for the population to adapt,
we translated the total effect size $x$ of each region 
to a probability of fixation
(by $p(x) \approx 2x/45$),
and then calculated $\sum_i x_i p(x_i)$.
If a region moves the trait an average of $x$ units towards the freshwater optimum
(averaging over dominance relationships),
Given the probability that an ``effect region'' fixes at a given trait value, 
$2 \times x / 45$ where $x$ is the effect size of a haplotype,
we were able to determine expected total effect sizes of the haplotypes that fix (detail about computation in methods).
Below, we see the mean of these expected total effect sizes across all lakes.
